---
title: "PAISAS Chapter 16"
output: html_notebook
---

# Functions

```{r}
library(tidyverse)
```


## 16.1 Making Some R Functions

A temperature in Celsius is converted to one in Kelvin by adding 273.15. (A temperature of  Celsius, 0 Kelvin, is the “absolute zero” temperature that nothing can be colder than.) Write a function called c_to_k that converts an input Celsius temperature to one in Kelvin, and test that it works.

```{r}
celsius_to_kelvin <- function(x){
  x + 273.15
}
```

```{r}
celsius_to_kelvin(0)
```
Write a function to convert a Fahrenheit temperature to Celsius. The way you do that is to subtract 32 and then multiply by $5/9$

```{r}
farenheit_to_celsius <- function(x){
  (x - 32) * 5/9
}
```

```{r}
farenheit_to_celsius(50)
```
```{r}
f_to_k <- function(x){
  c <- farenheit_to_celsius(x)
  k <- celsius_to_kelvin(c)
  return(k)
}
```


```{r}
f_to_k(32)
```
```{r}
far_kelvin <- function(x){
  celsius_to_kelvin(farenheit_to_celsius(x))
}
```

```{r}
far_kelvin(32)
```
```{r}
f_to_c <- function(F = 68) {
  (F - 32) * 5 / 9
}
```

```{r}
f_to_c(68)
f_to_c()
```
```{r}
temps <- seq(30, 80, 10)
temps
```
```{r}
f_to_c(temps)
```
```{r}
tibble(temps = seq(30, 80, 10)) %>%
  mutate(celsius = f_to_c(temps))
```

```{r}
temps <- seq(30, 80, 10)
temps %>%
  enframe(value = "fahrenheit") %>%
  mutate(celsius = f_to_c(temps))
```



