---
title: "PASIAS CHapter 9 - The sign test"
author: "Mohammad Amr Khan"
output: html_notebook
---

**Will only contains questions 1 - 3**

## Packages

```{r}
library(tidyverse)
```

## Running a maze

```{r}
myurl <- "http://www.utsc.utoronto.ca/~butler/c32/maze.txt"
times <- read_delim(myurl, " ")
```
```{r}
times
```

**Manual Sign test**

```{r}
times %>% count(time > 60)
```

```{r}
p <- sum(dbinom(0:5, 21, 0.5))
p
```

```{r}
sum(dbinom(16:21, 21, 0.5))
```
since we want to the two sided p val
```{r}
2*p
```

*Using SMMR*

```{r}
install.packages("devtools")
```

```{r}
library(devtools)
```
```{r}
install_github("nxskok/smmr")
```

```{r}
library(smmr)
```


```{r}
sign_test(times, time, 60)
```

two-sided is 0.0266 

```{r}
pval_sign(60, times, time)
```
Gives the two sided p-value

to get the confidence interval 
```{r}
ci_median(times, time)
```

## Chocolate Chips

A famous cookie manufacturer claims that their bags of chocolate chip cookies contain “more than 1100 chocolate chips on average”. A diligent group of students buys 16 bags of these cookies and counts the number of chocolate chips in each bag. The results are in http://www.utsc.utoronto.ca/~butler/c32/chips.txt.

Single column so any method of reading will work

```{r}
chip_url <- "https://www.utsc.utoronto.ca/~butler/c32/chips.txt"
chip_data <- read_table(chip_url)
```

```{r}
chip_data
```

```{r}
ggplot(chip_data, aes(x=chips)) + geom_histogram(bins = 4)
```

Right skew 

```{r}
chip_data %>% count(chips < 1100)
```

one less than 1100.

```{r}
chip_data %>% mutate(less=(chips<1100)) %>% count(less)
```

```{r}
sign_test(chip_data, chips, 1100)
```

We reject null hypothesis. 

```{r}
ci_median(chip_data, chips)
```
135 to 1325. I would round these off to whole numbers, since the data values are all whole numbers. These values are all above 1100, which supports the conclusion we got above that the median is above 1100. This is as it should be, because the CI is “all those medians that would not be rejected by the sign test”.

## The Power of the Sign Test


```{r}
power.t.test(n=10, delta = 50 -40, sd = 15, type = "one.sample")
```

Power of 50%.

```{r}
x=rnorm(10,50,15)  
x
```
```{r}
tibble(x) %>% count(x<40)
```

## 9.4 Sugar in Breakfast Cereals

The data is available here `http://www.utsc.utoronto.ca/~butler/c32/cereal-sugar.txt`
It is a text file, that looks to be delimited by spaces

```{r}
cereal_url <- "http://www.utsc.utoronto.ca/~butler/c32/cereal-sugar.txt"
cereal_data <- read_delim(cereal_url, delim = " ")
```

There are two columns in the data, one which contains text and another with 
numbers. 

This is expected

```{r}
cereal_data
```

```{r}
cereal_data %>% group_by(who) %>% 
  summarise(sugar_mean=mean(sugar))
```

The children cereal have a higher sugar mean than the adult brands.

```{r}
ggplot(cereal_data, aes(x=who, y=sugar)) + geom_boxplot()
```
The spread of the data is different since the boxplot are of different sizes
The outliers in adults pull the mean of the adult sugar level up.

A normal t-test is not reliable here since the t test assumes normality of the data, here the data is not normal so we should not be using the two sample t test

```{r}
median_test(cereal_data, sugar, who)
```

The p value is smaller than 0.05 so we conclude that the median sugar level in cereal is different

```{r}
t.test(sugar ~ who, data = cereal_data)
```

The p value is much smaller but we get the confidence intervals here

## Fear of math

Two new short courses have been proposed for helping students who suffer from severe math phobia. The courses are labelled A and B. Ten students were randomly allocated to one of these two courses, and each student’s score on a math phobia test was recorded after they completed their course. The math phobia test produces whole-number scores between 0 and 10, with a higher score indicating a greater fear of mathematics. The data can be found in link. We start with R for this question.

Data is available here `http://www.utsc.utoronto.ca/~butler/c32/mathphobia.txt`

Data is in a text file where each entry is separated by spaces so delimited 

```{r}
fear_url <- "http://www.utsc.utoronto.ca/~butler/c32/mathphobia.txt"
fear_data <- read_delim(fear_url, delim = " ")
```
Two columns one for course and another for phobis
which is what is expected


```{r}
fear_data
```

```{r}
fear_data %>% group_by(course) %>% 
  summarise(how_many=n())
```

```{r}
t.test(phobia ~ course, data=fear_data)
```

P value is 0.445 so we cannot reject the null hypothesis that there is no difference is mean phobia scores

```{r}
ggplot(fear_data, aes(x=course, y=phobia)) + geom_boxplot()
```

The spread of the data is different event though the medians looks to be similar.

The t test assumes normality and the boxplot shows evidence that the data is not normal.

```{r}
ggplot(fear_data, aes(sample=phobia)) + stat_qq() + stat_qq_line() + facet_wrap(~ course)
```

The points are close to the line, however there is a limited sample size so the central limit theorem does not apply. 
```{r}
median_test(fear_data, phobia, course)
```


P value is 0.659 so we cannot reject null median hypothesis

## Medical Instructions



